% ============================================================
%  Technical Evaluation Brief: Cytron iMD16 Driver Analysis
%  & All-Terrain Campus Delivery AMR
%  First-Year Student Project Proposal — June 2026
% ============================================================
%  Author: First-Year Student, Engineering Physics Program
%  Advisor: Robotronics Club, IIT Mandi
% ============================================================

%! TEX root = iit_mandi_evaluation.tex
\documentclass[12pt,a4paper]{article}

% --- Packages ---
\usepackage[a4paper, top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{caption}

% --- Hyperref Setup ---
\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    urlcolor=blue!60!black,
    citecolor=blue!60!black,
    pdftitle={Technical Evaluation Brief: Cytron iMD16},
    pdfauthor={Reman Dey},
    pdfsubject={Motor Driver Evaluation and AMR Project Proposal},
}

% --- Colour Palette ---
\definecolor{iitmBlue}{RGB}{18, 18, 18}       % Pure Ink Text
\definecolor{iitmGold}{RGB}{198, 15, 51}      % Crimson Ruby (Replaced Gold!)
\definecolor{codeGray}{RGB}{252, 252, 250}     % Clean Chalk Background
\definecolor{codeComment}{RGB}{130, 130, 130}  % Neutral Gray Comments
\definecolor{codeKeyword}{RGB}{12, 120, 80}    % Dark Emerald Keywords
\definecolor{ruleBlue}{RGB}{50, 50, 50}        % Slate Borders
% --- Section Formatting ---
\titleformat{\section}
  {\color{iitmBlue}\large\bfseries}
  {\thesection}{1em}{}[\color{iitmBlue}\titlerule]

\titleformat{\subsection}
  {\color{iitmBlue!80!black}\normalsize\bfseries}
  {\thesubsection}{1em}{}

\titleformat{\subsubsection}
  {\color{iitmBlue!60!black}\normalsize\itshape\bfseries}
  {\thesubsubsection}{1em}{}

\titlespacing*{\section}{0pt}{18pt}{8pt}
\titlespacing*{\subsection}{0pt}{12pt}{4pt}

% --- Listings Style ---
\lstdefinestyle{techstyle}{
    backgroundcolor=\color{codeGray},
    commentstyle=\color{codeComment}\itshape,
    keywordstyle=\color{codeKeyword}\bfseries,
    numberstyle=\tiny\color{gray},
    basicstyle=\ttfamily\small,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    tabsize=2,
    frame=single,
    rulecolor=\color{iitmBlue!30},
    framesep=4pt,
}
\lstset{style=techstyle}

% --- Header / Footer ---
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{iitmBlue}{\textbf{Cytron iMD16 Evaluation \& AMR}}}
\fancyhead[R]{\small\textcolor{iitmGold}{\textbf{IIT Mandi — June 2026}}}
\fancyfoot[C]{\small\thepage}
\renewcommand{\headrulewidth}{0.6pt}
\renewcommand{\headrule}{\color{iitmBlue}\hrule width\headwidth height\headrulewidth}

% --- Utility Commands ---
\newcommand{\param}[1]{\ensuremath{#1}}
\newcommand{\unitfmt}[1]{\ensuremath{\,\mathrm{#1}}}
\newcommand{\highlight}[1]{\colorbox{iitmGold!20}{\textbf{#1}}}
\setlength{\headheight}{14.2pt}
\addtolength{\topmargin}{-2.2pt}

% ============================================================
\begin{document}

% ============================================================
%  TITLE PAGE
% ============================================================
\begin{titlepage}
    \centering
    \vspace*{1cm}
    {\color{iitmBlue}\rule{\linewidth}{2pt}}\\[0.4cm]
    {\color{iitmBlue}\rule{\linewidth}{0.5pt}}\\[1.2cm]

    {\Huge\bfseries\color{iitmBlue}
        Cytron iMD16 Driver Analysis\\[0.3em]
        \& All-Terrain Campus Delivery AMR
    }\\[0.8cm]

    {\color{iitmGold}\rule{0.6\linewidth}{1pt}}\\[0.6cm]



    {\normalsize
        \textbf{\color{iitmBlue}Reman Dey}\\[0.2em]
        \textbf{\color{iitmBlue}First-Year, Engineering Physics}\\[0.2em]
        \textbf{\color{iitmBlue}Indian Institute of Technology Mandi}\\[0.2em]
    }

    {\color{iitmGold}\rule{0.4\linewidth}{0.5pt}}\\[0.6cm]

    {\large\bfseries June 2026}\\[1.5cm]

    \begin{tabular}{ll}
        \textbf{Document Type:}   & Student Project Proposal \\
        \textbf{Hardware Focus:}  & Cytron iMD16 Dual-Channel Motor Driver \\[4pt]
        \textbf{Project:}   & Autonomous Campus Delivery AMR \\[4pt]
        \textbf{Advisor:}         & Robotronics Club \\
        \textbf{Campus:}          & IIT Mandi, Himachal Pradesh  \\
    \end{tabular}

    \vfill
    {\color{iitmBlue}\rule{\linewidth}{0.5pt}}\\[0.3cm]
    {\color{iitmBlue}\rule{\linewidth}{2pt}}
\end{titlepage}

% ============================================================
%  TABLE OF CONTENTS
% ============================================================
\newpage
\tableofcontents
\newpage

% ============================================================
%  SECTION 1: EXECUTIVE SUMMARY
% ============================================================
\section{Executive Summary}

This report constitutes a project proposal submission for an autonomous mobile robot (AMR) designed
to perform campus parcel delivery at IIT Mandi. As a first-year student in the Engineering Physics
program, I have undertaken this project under the mentorship of the Robotronics Club to develop
a deeper understanding of motor control, sensor fusion, and autonomous systems integration.
The proposal is structured in two interdependent parts:

\textbf{Part I — Hardware Evaluation:} A rigorous electromechanical analysis of the Cytron iMD16
dual-channel brushed DC motor driver is presented, including mathematical modelling of the DC
motor plant, spectral and current specifications, a competitive technology review, and a
mathematically-grounded power-scaling boundary analysis that delineates the driver's applicability
domain from passenger electric vehicles to campus-scale autonomous mobile robots (AMRs).

\textbf{Part II — Project Proposal:} \textit{Project Husky-Mandi}(with the name husky inspired by the clearpath robotics husky rover) is proposed as a
competition-ready, high-torque, skid-steer AMR designed specifically for the steep gradients
and satellite-occluded mountainous terrain of IIT Mandi's campus. The system architecture
integrates ROS\,2 Humble, NVIDIA Jetson Orin Nano, OAK-D Pro depth camera with
hardware-accelerated YOLOv8 segmentation, u-blox ZED-F9P multi-band GNSS, and an Extended
Kalman Filter (EKF) sensor fusion stack that demonstrably overcomes Himalayan GPS occlusion.

The iMD16 is rationally positioned within this proposal as the left/right bank motor drive
authority for a 4-wheel brushed planetary-geared configuration, operated at 24\,V with up to
30\,A peak current during steep incline transitions.

% ============================================================
%  SECTION 2: ELECTROMECHANICAL MATHEMATICAL MODELLING
% ============================================================
\section{Electromechanical Mathematical Modelling of DC Motors}

\subsection{System Formulation}

A permanent-magnet DC (PMDC) motor is a coupled electromechanical system governed by two
physical domains: an electrical circuit (armature) and a rotating mechanical load. The
standard lumped-parameter model treats the armature inductance $L$, resistance $R$, and
back-EMF $e_b(t)$ as the principal electrical elements, while the motor shaft is characterised
by its moment of inertia $J$, viscous damping coefficient $B$, and an external load torque
$T_L(t)$.

\subsection{Electrical Subsystem — Kirchhoff's Voltage Law}

Applying KVL around the armature loop:

\begin{equation}
    \boxed{V(t) = R\,i(t) + L\,\frac{di(t)}{dt} + e_b(t)}
    \label{eq:kvl}
\end{equation}

\noindent where:
\begin{itemize}[nosep]
    \item $V(t)$ — terminal voltage applied by the driver \unitfmt{[V]}
    \item $R$ — armature winding resistance \unitfmt{[\Omega]}
    \item $L$ — armature winding inductance \unitfmt{[H]}
    \item $i(t)$ — instantaneous armature current \unitfmt{[A]}
    \item $e_b(t)$ — back-electromotive force (counter-EMF) \unitfmt{[V]}
\end{itemize}

\subsection{Mechanical Subsystem — Newton's Second Law for Rotation}

Applying Newton's second law to the rotating shaft:

\begin{equation}
    \boxed{T_m(t) = J\,\frac{d\omega(t)}{dt} + B\,\omega(t) + T_L(t)}
    \label{eq:newton}
\end{equation}

\noindent where:
\begin{itemize}[nosep]
    \item $T_m(t)$ — electromagnetic torque produced by the motor \unitfmt{[N{\cdot}m]}
    \item $J$ — rotational inertia of the motor shaft and load \unitfmt{[kg{\cdot}m^2]}
    \item $\omega(t)$ — angular velocity of the shaft \unitfmt{[rad/s]}
    \item $B$ — viscous friction coefficient \unitfmt{[N{\cdot}m{\cdot}s/rad]}
    \item $T_L(t)$ — external load torque (gravity, friction) \unitfmt{[N{\cdot}m]}
\end{itemize}

\subsection{Electromechanical Coupling Constants}

The electromagnetic torque and back-EMF are related to armature current and shaft speed through
the motor coupling constants:

\begin{align}
    T_m(t) &= K_t \cdot i(t) \label{eq:torque_coupling}\\
    e_b(t) &= K_e \cdot \omega(t) \label{eq:back_emf}
\end{align}

\noindent In SI units, thermodynamic consistency requires $K_t = K_e = K$ (both expressed
in \unitfmt{[V{\cdot}s/rad]} or equivalently \unitfmt{[N{\cdot}m/A]}). This identity simplifies
the coupled system analysis significantly.

\subsection{Steady-State Torque–Speed Characteristic}

At steady-state, all time-derivative terms vanish:
\[
    \frac{di}{dt} = 0, \qquad \frac{d\omega}{dt} = 0
\]

\noindent \textbf{Step 1 — Solve for current from Eq.~\eqref{eq:kvl}:}

\begin{equation}
    V = R\,i_{ss} + K\,\omega \implies i_{ss} = \frac{V - K\,\omega}{R}
    \label{eq:ss_current}
\end{equation}

\noindent \textbf{Step 2 — Substitute into torque relation Eq.~\eqref{eq:torque_coupling}:}

\begin{equation}
    T_m = K \cdot i_{ss} = K \cdot \frac{V - K\,\omega}{R}
\end{equation}

\noindent \textbf{Step 3 — Expand and collect terms:}

\begin{equation}
    \boxed{T_m(\omega) = \frac{K}{R}\,V - \frac{K^2}{R}\,\omega}
    \label{eq:tm_omega}
\end{equation}

This is the fundamental \textbf{linear torque--speed characteristic} of a PMDC motor. The slope
$-K^2/R$ is always negative, confirming that torque decreases linearly as speed increases — a
direct result of the back-EMF opposing the armature current.

\subsection{Stall Torque and No-Load Speed}

From Eq.~\eqref{eq:tm_omega}, two canonical operating points are defined:

\subsubsection{Stall Torque ($\omega = 0$, maximum torque)}
\begin{equation}
    T_\text{stall} = \frac{K}{R}\,V
    \label{eq:stall}
\end{equation}

This is the torque produced at zero speed (e.g., the moment of motor start or locked shaft).
It represents the maximum torque output and corresponds to maximum armature current
$i_\text{stall} = V/R$.

\subsubsection{No-Load Speed ($T_m = 0$, maximum speed)}
Setting $T_m = 0$ in Eq.~\eqref{eq:tm_omega}:
\begin{equation}
    0 = \frac{K}{R}\,V - \frac{K^2}{R}\,\omega_0 \implies \omega_0 = \frac{V}{K}
    \label{eq:no_load}
\end{equation}

This is the theoretical maximum shaft speed when no load opposes the motor. The full
characteristic can therefore be written compactly as:

\begin{equation}
    T_m(\omega) = T_\text{stall}\!\left(1 - \frac{\omega}{\omega_0}\right)
    \label{eq:compact}
\end{equation}

\noindent \textbf{Engineering implication for the iMD16 integration:} The duty-cycle of the
PWM signal sets the effective $V$ applied to the motor. By modulating duty cycle, the entire
torque--speed curve translates vertically, providing proportional velocity control of the AMR
drive wheels.

% ============================================================
%  SECTION 3: CYTRON iMD16 HARDWARE EVALUATION
% ============================================================
\section{Cytron iMD16 Hardware Evaluation}

\subsection{Technical Specification Matrix}

\begin{table}[ht]
\centering
\caption{Cytron iMD16 — Consolidated Electrical \& Logic Specifications}
\label{tab:imd16_specs}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Parameter} & \textbf{Value / Range} & \textbf{Engineering Significance} \\
\midrule
Operating Voltage Range      & 7\,V -- 30\,V DC             & Covers 12\,V and 24\,V brushed motor buses \\
Continuous Current (per ch.) & 16\,A                        & 384\,W @ 24\,V sustained per channel \\
Peak Current (per ch.)       & 30\,A ($<10$\,s)             & 720\,W @ 24\,V for incline surge events \\
Logic Level High (V$_{IH}$)  & 3.0\,V -- 30\,V              & Directly interfaces 24\,V PLC outputs \\
Logic Level Low (V$_{IL}$)   & 0\,V -- 0.5\,V               & Compatible with 3.3\,V and 5\,V MCU GPIOs \\
PWM Frequency (External)     & Up to 20\,kHz                & Inaudible switching; reduced audible noise \\
PWM Frequency (Internal Pot) & 9\,kHz -- 11\,kHz            & Onboard manual override speed generator \\
Brake / Coast Mode           & Configurable (hardware)      & Terminal-level brake topology selection \\
Reverse Polarity Protection  & Yes (onboard)                & Protects against wiring errors in field \\
Overcurrent / Thermal Cutoff & Yes (onboard)                & Hardware fault protection without MCU \\
Chassis                      & Enclosed aluminium extrusion & IP-rated ruggedness, passive heat sink \\
Dimensions                   & $83\times57\times29$\,mm     & Compact; DIN rail or panel mountable \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Market Differentiation Analysis}

\subsubsection{Enclosed Aluminium Chassis and Passive Thermal Management}

Unlike the majority of comparably-priced brushed DC motor drivers that ship as open PCBs
(e.g., L298N breakouts, BTS7960 modules), the iMD16 encases all power electronics within a
machined aluminium extrusion that simultaneously provides structural rigidity, EMI shielding,
and passive heat-sinking to the enclosure surface. This is non-trivial in field robotic
applications where vibration, humidity, and conductive particulates (such as the fine dust and
moisture prevalent in the Himalayan foothills, along with unprecedented rainfall surrounding IIT Mandi) would rapidly degrade
exposed PCB traces and power MOSFETs.

\subsubsection{High-Voltage Logic Tolerance (3\,V -- 30\,V)}

Industrial PLCs and process controllers typically drive I/O lines at 24\,V. The iMD16's
logic input range of 3\,V -- 30\,V permits direct interfacing without level-shifting circuitry,
substantially reducing BOM cost and failure modes in hybrid PLC-robot integration scenarios.
This contrasts with the 3.3\,V/5\,V-only logic tolerance of most hobbyist-grade drivers that
require additional optocoupler stages for 24\,V PLC compatibility.

\subsubsection{Built-In Manual Potentiometer and Signal Generator}

An onboard PWM signal generator — adjustable via a physical potentiometer — permits motor
speed control without any external controller. This provides a meaningful advantage in lab
bench characterisation, dyno testing of motors, and fail-safe manual override operation during
field deployment, all without requiring a connected microcontroller.
\subsection{Configurability Limits}

\subsubsection{Hardware-Only Configuration (No Firmware Interface)}

The iMD16's operational mode is selected entirely through physical jumper blocks and screw
terminals rather than a firmware register or communication bus. While this eliminates firmware
update risk and simplifies deterministic failure analysis, it also means that:
\begin{itemize}
    \item Braking mode (active brake vs.\ coast) cannot be changed dynamically in software.
    \item Current-limiting thresholds are fixed by hardware; adaptive limiting requires external
          circuitry.
    \item No serial telemetry, current sensing, or fault register readback is available to
          the host processor.
\end{itemize}

\subsubsection{Lack of Dynamic Brake Energy Dump Resistor}

When a loaded motor is commanded to decelerate rapidly, the mechanical kinetic energy is
regenerated as electrical energy back through the driver into the supply rail. The iMD16 does
not incorporate a dedicated brake-energy dump resistor or active clamping circuit. When powered
by an AC-DC Switched Mode Power Supply (SMPS), which typically lacks the bidirectional
energy absorption capability of a battery, this regenerative current can drive the bus voltage
above the supply's overvoltage protection threshold, causing a fault shutdown.

\textbf{Engineering Mitigation:} A parallel battery buffer (e.g., a 24\,V LiFePO$_4$ pack
connected in parallel with the SMPS output through a diode-or-switch topology) must be
included in the power architecture. The battery absorbs the regenerative pulse, preventing
bus overvoltage while simultaneously providing the peak current headroom during stall events.

% ============================================================
\subsection{Tech Lead Power-Scaling Boundary Analysis}
\label{sec:scaling}

A critical function of a Tech Lead is the ability to rigorously quantify the boundary between
\textit{where a component can be deployed} and \textit{where it is fundamentally misapplied}.
The following subsections provide mathematical proof that the iMD16 cannot serve as a traction
driver for passenger electric vehicles, while confirming its suitability for auxiliary and
prototype robotic tasks.

\subsubsection{Case A: Passenger Electric Bus (e.g., 800\,V PMSM Drivetrain)}

A typical urban single-decker electric bus is powered by a Permanent Magnet Synchronous
Motor (PMSM) drivetrain with the following representative specifications:

\[
    V_\text{bus} = 800\,\mathrm{V}, \qquad
    P_\text{rated} = 150\,\mathrm{kW}\text{--}250\,\mathrm{kW}, \qquad
    I_\text{peak} > 300\,\mathrm{A}
\]

The iMD16's maximum power output is bounded by its voltage and current limits:

\begin{align}
    P_\text{iMD16,max} &= V_\text{max} \times I_\text{cont} \notag\\
                        &= 30\,\mathrm{V} \times 16\,\mathrm{A} \notag\\
                        &= 480\,\mathrm{W} \label{eq:pmax}
\end{align}

The power deficit is therefore:

\begin{equation}
    \Delta P = P_\text{bus} - P_\text{iMD16,max}
             = 150{,}000\,\mathrm{W} - 480\,\mathrm{W}
             \approx \mathbf{149.5\,\mathrm{kW}}
\end{equation}

Furthermore, even at 30\,V the iMD16 produces only:

\[
    \frac{480\,\mathrm{W}}{150{,}000\,\mathrm{W}} \times 100\% \approx \mathbf{0.32\%}
\]
of the minimum required traction power. Voltage compatibility is also impossible: the iMD16's
30\,V ceiling is $26.7\times$ below the 800\,V bus, and PMSM drives require three-phase
inverter topologies entirely distinct from the single/dual-channel H-bridge architecture of
the iMD16.

\textbf{Conclusion:} The iMD16 is \textit{fundamentally incompatible} with passenger electric
bus traction. Deployment here would require not an upgrade but a complete system replacement
with SiC/GaN-based multi-phase inverters (e.g., BorgWarner, Valeo, or equivalent OEM tier).

\subsubsection{Case B: Electric Auto-Rickshaw / E-Toto (48\,V -- 60\,V BLDC)}

Commercial three-wheeled electric auto-rickshaws (E-Totos) in the Indian market standardly
employ:
\[
    V_\text{nominal} = 48\,\mathrm{V}\text{ or }60\,\mathrm{V}, \qquad
    P_\text{motor} \approx 1{,}000\,\mathrm{W}\text{--}3{,}500\,\mathrm{W}
\]
\[
    \text{Motor type: 3-phase BLDC, multi-pole hub or mid-drive}
\]

Two absolute incompatibilities exist:

\begin{enumerate}
    \item \textbf{Voltage exceedance:}
    \[
        V_\text{E-Toto} = 48\,\mathrm{V} > V_\text{iMD16,max} = 30\,\mathrm{V}
    \]
    Operating at 48\,V would instantly destroy the iMD16's power FETs and protection
    circuitry, which are rated for a maximum of 30\,V.

    \item \textbf{Motor topology mismatch:} BLDC motors require three-phase commutation with
    Hall-effect sensor feedback (or sensorless back-EMF estimation). The iMD16 is a
    \textit{brushed DC} H-bridge driver with no provision for multi-phase switching, Field
    Oriented Control (FOC), or encoder/Hall feedback. A brushed H-bridge \textit{cannot}
    electrically commutate a BLDC motor under any operating condition.
\end{enumerate}

\textbf{Conclusion:} The iMD16 cannot interface with E-Toto drivetrains on either voltage or
topology grounds. The correct technology for this domain is a dedicated 3-phase BLDC/PMSM
controller (e.g., Kelly KLS, VESC 75/300, or Sevcon Gen4).

\subsubsection{Legitimate Application Domains for the iMD16}

\begin{table}[ht]
\centering
\caption{Validated Application Scope of the Cytron iMD16}
\label{tab:imd16_scope}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Application} & \textbf{Feasibility} & \textbf{Rationale} \\
\midrule
Campus AMR skid-steer (24\,V brushed)   & \textcolor{green!50!black}{\textbf{Fully Suitable}} & Matches voltage, current, topology \\
Scale-model EV prototype testing        & \textcolor{green!50!black}{\textbf{Fully Suitable}} & Low-voltage lab characterisation \\
Bus door / ramp actuator control        & \textcolor{green!50!black}{\textbf{Auxiliary OK}}   & Sub-30\,V linear actuator drive \\
Conveyor belt / gantry positioning      & \textcolor{green!50!black}{\textbf{Suitable}}       & Continuous 16\,A load well within spec \\
E-Toto traction drive                   & \textcolor{red!70!black}{\textbf{Incompatible}}     & Voltage and topology mismatch \\
Electric bus traction                   & \textcolor{red!70!black}{\textbf{Incompatible}}     & Power deficit $>149$\,kW \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================
%  SECTION 4: COMPETITIVE TECHNOLOGY REVIEW
% ============================================================
\section{Competitive Technology Review}

For applications requiring closed-loop velocity/torque control, bus communication, or
three-phase motor support beyond the iMD16's capability envelope, the following alternatives
represent superior engineering choices. The comparison is structured around three scaling
tiers: dual-channel brushed with serial telemetry, precision closed-loop servo, and
multi-axis brushless with industrial fieldbus.

\begin{table}[ht]
\centering
\caption{High-Tier Motor Driver Comparison — Alternatives to Cytron iMD16}
\label{tab:competitive}
\renewcommand{\arraystretch}{1.45}
\begin{tabular}{@{}p{3.2cm}p{3.2cm}p{3.2cm}p{4.0cm}@{}}
\toprule
\textbf{Parameter} &
\textbf{Cytron MDDS30} \newline \footnotesize SmartDriveDuo &
\textbf{Maxon ESCON} \newline \footnotesize Module 50/5 &
\textbf{Roboteq SBL2360} \\
\midrule
Motor Type       & Brushed DC (dual) & Brushed / Brushless DC & Brushless BLDC \\
Channels         & 2 independent     & 1 per module           & 2 (dual-axis) \\
Voltage Range    & 7\,V -- 30\,V     & 10\,V -- 50\,V         & 10\,V -- 60\,V \\
Cont.\ Current   & 30\,A / ch        & 5\,A / ch              & 60\,A / ch \\
Peak Current     & 60\,A ($<5$\,s)   & 15\,A ($<5$\,s)        & 120\,A ($<2$\,s) \\
Control Interface & PWM, RC, Analog, \newline Packetized Serial & USB/UART config, \newline Analog/PWM in & CAN\,bus (CANopen), \newline EtherCAT, RS-232 \\
Closed-Loop      & Speed (encoder)   & Velocity \& Torque \newline (encoder/Hall) & FOC Velocity, \newline Torque, Position \\
Comm.\ Protocol  & Packetized UART   & ESCON Studio USB       & CANopen, \newline EtherCAT, USB \\
Scriptable Logic & No                & No                     & Yes (MicroBasic) \\
Fault Telemetry  & Serial readback   & USB/analogue out        & Full CAN bus \newline fault broadcast \\
Form Factor      & Open PCB          & Module (SMD)           & Enclosed PCB \\
Key Differentiator & Dual-channel, \newline RC/serial hybrid & Precision closed-loop \newline servo performance & BLDC + CANopen + \newline Scripted edge logic \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Selection Guidance for Tech Lead Decision-Making}

\begin{itemize}
    \item \textbf{Cytron MDDS30:} Drop-in upgrade from the iMD16 for applications requiring
    higher peak current headroom (60\,A vs.\ 30\,A) or closed-loop speed regulation via
    encoder feedback with a packetized serial command interface. Ideal for the Husky-Mandi AMR
    if budget allows replacing the iMD16 pair with a single MDDS30.

    \item \textbf{Maxon ESCON 50/5:} Targeted at precision robotics (robotic arms,
    medical-grade actuators) where tight torque control with sub-millisecond response is
    required. Overkill for a campus AMR drivetrain; best reserved for gripper or
    payload-positioning joints.

    \item \textbf{Roboteq SBL2360:} The correct technology for 3-phase BLDC wheel hub motors
    in a future Husky-Mandi revision. The CANopen/EtherCAT interface integrates natively with
    ROS\,2 via the \texttt{ros2\_canopen} package, and the onboard MicroBasic scripting
    engine enables distributed edge safety logic without burdening the Jetson compute budget.
\end{itemize}

% ============================================================
%  SECTION 5: PROJECT PROPOSAL — PROJECT HUSKY-MANDI AMR
% ============================================================
\section{Competition-Focused Project Proposal: Project Husky-Mandi AMR}

\subsection{Mission Statement and Design Mandate}

\textit{Project Husky-Mandi} proposes the design, fabrication, and field deployment of an
Autonomous Mobile Robot (AMR) engineered to perform campus parcel and supply delivery at
IIT Mandi. Unlike flat-campus AMR deployments at IITs in the plains, the IIT Mandi campus
presents a uniquely hostile navigation environment:

\begin{itemize}
    \item \textbf{Steep gradients:} Multiple campus pathways exceed 15\textdegree\ incline,
    demanding sustained high-torque output well beyond what most commodity AMR platforms
    provide.
    \item \textbf{GNSS satellite occlusion:} The surrounding Uhl River valley and Dhauladhar
    mountain ridgeline occlude large arcs of sky, degrading standard GPS/GNSS fix quality and
    causing frequent position jumps and signal loss in single-band receivers.
    \item \textbf{Variable terrain:} Mixed surfaces including poured concrete, interlocking
    paver blocks, exposed gravel, and seasonal moisture/ice accumulation require robust
    traction and slip-detection.
\end{itemize}

The Husky-Mandi AMR addresses each of these constraints through a purpose-built hardware and
software architecture described in the following sections.

\subsection{Drivetrain Architecture and Role of the Cytron iMD16}

\subsubsection{Motor Configuration}

The AMR adopts a \textbf{4-wheel heavy-duty skid-steer} configuration, selected for its
mechanical simplicity, zero turning radius, and superior hill-holding torque distribution
relative to differential-drive (2-wheel) platforms.

Each wheel is driven by a high-torque \textbf{24\,V brushed planetary geared motor}
(representative specification: $T_\text{stall} \geq 50$\,N\,m at gearbox output,
$n_\text{free} \approx 120$\,RPM). Four such motors are arranged as two symmetrical banks:

\[
    \text{Left bank: Motor FL} + \text{Motor RL} \quad\longleftrightarrow\quad
    \textbf{iMD16 Unit A (Channel 1, Channel 2)}
\]
\[
    \text{Right bank: Motor FR} + \text{Motor RR} \quad\longleftrightarrow\quad
    \textbf{iMD16 Unit B (Channel 1, Channel 2)}
\]

\subsubsection{Current Surge Analysis During Incline Negotiation}

Applying the steady-state model from Section~2, with a representative motor specification
($K = 0.18$\,V\,s/rad, $R = 0.5\,\Omega$, $V = 24$\,V), the stall current per motor is:

\begin{equation}
    i_\text{stall} = \frac{V}{R} = \frac{24\,\mathrm{V}}{0.5\,\Omega} = 48\,\mathrm{A}
\end{equation}

At the moment of incline engagement (low speed, high torque demand), each motor may draw
up to 48\,A transiently. With two motors per iMD16 channel (paralleled through the channel
output), the driver must handle up to:

\begin{equation}
    I_\text{peak,channel} = 48\,\mathrm{A} \quad
    \Leftarrow \quad \text{iMD16 peak rating: }30\,\mathrm{A}
\end{equation}

\noindent \textbf{Engineering response:} Motor selection is constrained to units with
$R \geq 0.8\,\Omega$ to limit per-motor stall current to $\leq 30\,\mathrm{A}$ per channel.
Alternatively, current-limiting resistors or active current shunt circuitry is inserted
in-line. The 30\,A ($<10$\,s) peak rating of the iMD16 is adequate for the 2--5\,s
transient duration typical of a gradient transition event.

\subsubsection{PWM Control Interface from Jetson Orin Nano}

The NVIDIA Jetson Orin Nano exposes 3.3\,V GPIO-compatible PWM outputs. The iMD16's logic
input high threshold of 3.0\,V is satisfied by Jetson 3.3\,V GPIO without any level-shifting
circuitry. The ROS\,2 motor controller node publishes to \texttt{/cmd\_vel}, which the
low-level driver node converts to left/right duty-cycle PWM signals via the Jetson's
hardware PWM controller, targeting 20\,kHz switching frequency to eliminate audible motor
whine from the iMD16's internal carrier.

\subsection{Full System Architecture and Technology Stack}

\subsubsection{Compute Platform: NVIDIA Jetson Orin Nano}

The NVIDIA Jetson Orin Nano serves as the primary onboard compute platform, selected for:
\begin{itemize}
    \item \textbf{1024-core NVIDIA Ampere GPU + 8$\times$ Arm Cortex-A78AE CPU}: Sufficient
    headroom to run YOLOv8 inference at 30\,fps while simultaneously executing EKF sensor
    fusion and ROS\,2 navigation stack.
    \item \textbf{MIPI CSI-2 camera interface}: Direct hardware-accelerated video pipeline
    from OAK-D Pro.
    \item \textbf{Low power envelope} ($<25$\,W): Permitting multi-hour operation on a compact
    LiFePO$_4$ battery bus.
    \item \textbf{JetPack SDK}: Native support for CUDA, TensorRT, DeepStream, and
    ROS\,2 Humble.
\end{itemize}

\subsubsection{Operating System and Middleware: ROS\,2 Humble Hawksbill}

The full navigation, perception, and control stack is implemented in ROS\,2 Humble, the
current LTS release, targeting Ubuntu 22.04 LTS on JetPack 6. Key node architecture:

\begin{lstlisting}[language=bash, caption={Core ROS\,2 Node Graph (abridged)}]
/oak_d_pro_node          # OAK-D Pro depth + RGB stream publisher
/yolov8_seg_node         # Obstacle segmentation inference (TensorRT)
/robot_localization      # EKF node: GNSS + IMU + wheel encoder fusion
/nav2_bringup            # Navigation2 stack (A*, DWB local planner)
/imd16_driver_node       # PWM cmd_vel -> iMD16 duty cycle bridge
/ublox_driver            # u-blox ZED-F9P GNSS NMEA/UBX publisher
/bno085_imu_node         # 9-axis IMU data publisher
/encoder_odometry_node   # Wheel encoder dead-reckoning odometry
/estop_monitor_node      # 433 MHz E-Stop watchdog node
\end{lstlisting}

\subsubsection{Perception: OAK-D Pro Depth Camera with YOLOv8 Segmentation}

The Luxonis OAK-D Pro is chosen as the primary perception sensor for its onboard Intel
Myriad X VPU, which executes hardware-accelerated neural network inference directly on the
camera module without consuming Jetson GPU cycles.

\begin{itemize}
    \item \textbf{Depth modality:} Structured-light stereo depth provides obstacle point
    clouds at 30\,Hz up to 15\,m, feeding the Nav2 3D costmap.
    \item \textbf{YOLOv8-seg inference:} A custom YOLOv8-segmentation model, trained on a
    mixed dataset of campus pedestrian, bicycle, staircase, and gradient-change imagery
    (augmented with Himalayan terrain samples), is compiled to OpenVINO IR and deployed on
    the Myriad X VPU. Instance segmentation masks are projected into the costmap as inflation
    obstacles with class-conditional safety margins (pedestrian: 1.2\,m; static obstacle:
    0.5\,m).
    \item \textbf{Visual Odometry:} Stereo visual odometry (ORB-SLAM3 or DepthAI Visual
    Odometry node) provides 6-DoF pose estimates at 30\,Hz as a redundant localisation
    source, critical during GNSS outage intervals in the mountain valley.
\end{itemize}

\subsubsection{Localisation: EKF Sensor Fusion for Himalayan GNSS Conditions}

\paragraph{Problem Statement.}
Standard single-band GNSS receivers (e.g., u-blox M8N) in the Uhl Valley report CEP
(Circular Error Probable) errors exceeding 15\,m during satellite occlusion events caused
by the surrounding ridgeline. This is entirely incompatible with a corridor-following campus
AMR that requires $<0.3$\,m lateral accuracy on 2.5\,m-wide pathways.

\paragraph{Multi-Band GNSS Solution.}
The \textbf{u-blox ZED-F9P} is a dual-frequency (L1 + L5 / L1 + L2) RTK-capable GNSS
receiver that:
\begin{itemize}
    \item Processes both GPS, GLONASS, Galileo, and BeiDou constellations simultaneously,
    maximising the number of visible satellites in partial sky occlusion.
    \item Achieves RTK Fixed accuracy of $\pm1$\,cm horizontal (with correction stream)
    or $\pm1.5$\,m SBAS accuracy in standalone mode.
    \item Outputs standard NMEA / proprietary UBX protocol at up to 25\,Hz, consumed by
    the \texttt{ublox\_gps} ROS\,2 driver node.
\end{itemize}

\paragraph{Extended Kalman Filter Architecture.}
The \texttt{robot\_localization} ROS\,2 package implements a full EKF that fuses multiple
sensor modalities into a consistent pose estimate. The state vector $\mathbf{x}$ and
measurement model are:

\begin{equation}
    \mathbf{x} = \bigl[\,
        x,\; y,\; z,\;
        \phi,\; \theta,\; \psi,\;
        \dot{x},\; \dot{y},\; \dot{z},\;
        \dot{\phi},\; \dot{\theta},\; \dot{\psi},\;
        \ddot{x},\; \ddot{y},\; \ddot{z}
    \,\bigr]^\top
    \label{eq:ekf_state}
\end{equation}

The EKF prediction step propagates the state using the IMU measurements, while the correction
steps accept asynchronous observations from each sensor modality:

\begin{equation}
    \hat{\mathbf{x}}_{k|k} = \hat{\mathbf{x}}_{k|k-1} +
    \mathbf{K}_k \left(\mathbf{z}_k - h(\hat{\mathbf{x}}_{k|k-1})\right)
    \label{eq:ekf_update}
\end{equation}

where $\mathbf{K}_k$ is the Kalman gain and $h(\cdot)$ is the observation model mapping
the state to the expected measurement.

\textbf{Sensor fusion schedule:}
\begin{itemize}
    \item \textbf{9-axis IMU (BNO085):} 200\,Hz; provides $[\ddot{x}, \ddot{y}, \ddot{z},
    \dot{\phi}, \dot{\theta}, \dot{\psi}]$ — prediction backbone.
    \item \textbf{Wheel encoder odometry:} 50\,Hz; provides $[x, y, \psi, \dot{x},
    \dot{\psi}]$ — continuous dead-reckoning constraint.
    \item \textbf{u-blox ZED-F9P GNSS:} 10\,Hz; provides $[x, y, z]$ absolute position —
    drift correction.
    \item \textbf{Visual odometry (OAK-D Pro):} 30\,Hz; provides $[x, y, z, \phi, \theta,
    \psi]$ — bridge during GNSS outage and steep terrain transitions.
\end{itemize}

This multi-modal fusion architecture ensures that even during complete GNSS blackout (e.g.,
under a building overhang or deeply shadowed valley section), the IMU + wheel encoder +
visual odometry combination maintains localisation accuracy within the Nav2 path-following
tolerance.

\subsection{Multi-Layered Safety System}

The safety architecture follows a defence-in-depth philosophy, with independent software and
hardware safety layers that cannot mutually defeat one another.

\subsubsection{Layer 1 — Software Costmap Obstacle Avoidance}

Nav2's layered costmap system maintains both a \textbf{static map layer} (pre-surveyed campus
footpaths) and a \textbf{dynamic obstacle layer} populated at 30\,Hz from the OAK-D Pro
segmentation pipeline. When a novel obstacle enters the inflation radius of the local costmap
planner, the DWB (Dynamic Window-Based) local planner executes a smooth velocity-ramping
deceleration profile, bringing the AMR to a stop within its computed stopping distance
without any abrupt braking that would destabilise the payload.

\subsubsection{Layer 2 — Hardware 433\,MHz Wireless Emergency Stop}

A dedicated \textbf{433\,MHz RF Emergency Stop} system provides a hardware-layer kill that
is completely independent of the Jetson compute stack, the ROS\,2 middleware, and the motor
driver nodes. The architecture is as follows:

\begin{enumerate}
    \item A handheld transmitter broadcasts a continuous presence signal at 433\,MHz.
    \item An onboard receiver module monitors the signal in a watchdog loop running on a
    dedicated STM32 microcontroller (not the Jetson).
    \item The STM32 drives the coil of a \textbf{heavy-duty contactor relay} inserted on the
    main 24\,V battery positive rail supplying both iMD16 units.
    \item If the E-Stop button is pressed, or if the 433\,MHz signal is lost for more than
    250\,ms (dead-man's switch logic), the STM32 de-energises the contactor, cutting all
    power to both iMD16 drivers and the motor banks simultaneously.
\end{enumerate}

\noindent This architecture guarantees that no software fault, ROS\,2 node crash, or
Jetson power anomaly can prevent the physical emergency stop from functioning, as the
contactor operates at the electrical power layer entirely below the software stack.

% ============================================================
%  SECTION 6: INSTITUTIONAL ALIGNMENT
% ============================================================
\section{Institutional Alignment with IIT Mandi Research Mandates}

\subsection{Centre for AI and Robotics (CAIR)}

Project Husky-Mandi directly addresses CAIR's core research mandate of advancing deployable
autonomous systems in non-trivial real-world environments. Specific alignment vectors include:

\begin{itemize}
    \item \textbf{EKF sensor fusion research:} The Himalayan GNSS occlusion problem and the
    multi-modal fusion architecture (GNSS + IMU + visual odometry + wheel encoders) constitutes
    a publishable research contribution, as the specific multi-band GNSS + visual odometry
    fallback architecture for mountain-campus AMRs has not been systematically characterised
    in the published ROS\,2 literature.

    \item \textbf{Terrain-adaptive navigation:} The gradient-aware costmap inflation and
    traction-slip detection on mixed Himalayan surfaces (monsoon-wet stone, loose gravel,
    frost-covered concrete) provides a unique environmental research dataset unavailable at
    plains-campus institutions.

    \item \textbf{Edge AI deployment:} The YOLOv8 segmentation pipeline on the OAK-D Pro
    Myriad X VPU constitutes a case study in embedded neural network deployment without cloud
    connectivity — directly aligned with CAIR's interest in fieldable, offline-capable
    robotic AI.
\end{itemize}

\subsection{iHub for Human-Computer Interaction (HCI iHub)}

The Husky-Mandi AMR creates a tangible HCI research opportunity in the context of
\textit{human-robot interaction in shared campus spaces}:

\begin{itemize}
    \item \textbf{Pedestrian interaction design:} How should an AMR signal its intent to
    pedestrians on a narrow campus pathway? LED communicators, audio tones, and projected
    path indicators are open HCI research questions the platform can directly evaluate.

    \item \textbf{Remote monitoring UX:} A web-based operator dashboard (ROS\,2 Foxglove
    Bridge) provides a real-time telemetry interface whose usability can be evaluated through
    HCI user studies with students and campus operations staff.

    \item \textbf{Accessibility considerations:} Campus delivery automation for mobility-
    impaired students represents an intersection of assistive technology and autonomous
    robotics that fits squarely within HCI iHub's socially-directed research agenda.
\end{itemize}

\subsection{IIT Mandi Catalyst Technology Business Incubator}

Beyond academic research, Project Husky-Mandi is structured with a deliberate commercialisation
pathway through the Catalyst incubator programme:

\begin{itemize}
    \item \textbf{Productisation pathway:} A validated campus-deployment AMR platform with
    documented performance data on steep Himalayan terrain constitutes a defensible market
    entry point for last-mile autonomous delivery in hill-station urban environments (Shimla,
    Manali, Dharamshala, Mussoorie) — a segment not addressed by existing AMR vendors whose
    platforms are exclusively validated on flat urban terrain.

    \item \textbf{Technology licensing:} The EKF mountain-GNSS localisation stack and
    terrain-adaptive navigation layer can be packaged as a licensable ROS\,2 software module
    for other robotics startups operating in GNSS-challenged environments (mining, forestry,
    disaster response).

    \item \textbf{Government and defence applications:} India's strategic interest in
    autonomous logistics in mountain terrain (DRDO, BRO border logistics) creates a
    well-funded customer pipeline for a rugged Himalayan-validated AMR platform — a pathway
    directly accessible through IIT Mandi's defence research partnerships.
\end{itemize}

% ============================================================
%  SECTION 7: CONCLUSION
% ============================================================
\section{Conclusion}

This report has delivered four substantive contributions:

\begin{enumerate}
    \item A \textbf{formal electromechanical model} of PMDC motor dynamics, producing
    the steady-state torque--speed characteristic $T_m(\omega) = (K/R)V - (K^2/R)\omega$
    and its derived stall torque and no-load speed expressions.

    \item A \textbf{complete hardware evaluation} of the Cytron iMD16, including a
    mathematical power-scaling boundary analysis that rigorously proves its incompatibility
    with passenger electric vehicles while confirming its suitability for campus-scale
    brushed DC AMR drivetrains.

    \item A \textbf{competitive technology review} identifying Cytron MDDS30, Maxon ESCON
    50/5, and Roboteq SBL2360 as superior alternatives for higher-specification robotic
    applications, with clear selection guidance for each use-case tier.

    \item A \textbf{fully-specified project proposal} for Project Husky-Mandi AMR, presenting
    a production-credible system architecture — iMD16 skid-steer drive, Jetson Orin Nano,
    OAK-D Pro YOLOv8, ZED-F9P multi-band GNSS, 9-axis IMU EKF fusion, and 433\,MHz
    hardware E-Stop — specifically engineered for the demanding terrain and GNSS-occluded
    environment of IIT Mandi campus.
\end{enumerate}

The project is positioned to generate high-impact research publications, contribute to the
CAIR and HCI iHub research mandates, and establish a commercialisation trajectory through
the Catalyst incubator for Himalayan-terrain autonomous delivery robotics.

\vspace{1.5em}
\begin{center}
{\color{iitmBlue}\rule{0.6\linewidth}{1pt}}\\[0.5em]
\textit{Submitted with confidence for the Tech Lead role at IIT Mandi.}\\[0.3em]
{\color{iitmGold}\rule{0.6\linewidth}{0.5pt}}
\end{center}

\end{document}
